### 2022-01-13 manuscript/manuscript.Rmd
Update manuscript.Rmd

abstract: |
  Critical scrutiny of published scientific research, often called post-publication peer review (PPPR), can highlight errors, limitations, or alternative interpretations that were not adequately addressed during traditional peer review. Currently, there is little empirical data documenting how academic journals handle this important aspect of scientific self-correction. We conducted a systematic and cross-disciplinary assessment of PPPR policies and practice amongst 330 influential journals operating across 22 scientific disciplines. 123 (37%) journals did not offer any options for submitting PPPR. 207 (63%) journals offered PPPR options such as letters, commentaries, or web comments, but often imposed limits on length (median 1000 words) and time-to-submit (median 12 weeks). The most restrictive limits were 175 words and 2 weeks. Amongst PPPR-offering journals, we found that 1.9% (95% CI [1.4, 2.6]) of a random sample of 2066 research articles published in 2018 were the subject of at least one PPPR. A detailed examination of 58 PPPRs found, for example, that the majority (76%) received an author reply. The majority of those replies (93%) asserted that the authors’ original claims were unchanged. Clinical Medicine stood out as the domain with the most active PPPR culture, with journals most frequently providing PPPR options and publishing the most PPPR, but also imposing the strictest length and time-to-submit limits. Overall, our findings suggest that influential academic journals often pose serious barriers to the cultivation, documentation, and dissemination of PPPR. We propose several policy changes that journals could adopt to promote a richer culture of post-publication critique.

@@

Table 1 shows how often journal policies imposed limits on PPPR in terms of length, time-to-submit, or number of references. Limits were sometimes expressed quantitatively, but sometimes they were qualitative and more ambiguous, for example, stating that PPPR should be ‘concise’ or address ‘recently published’ articles. Often there was no information at all about a particular limit. Occasionally policies explicitly asserted that there was no limit. This happened `r d_policy %>% filter(wordLimits == 'NO') %>% nrow()` time for length limits, `r d_policy %>% filter(timeLimits == 'NO') %>% nrow()` times for time-to-submit limits, and `r d_policy %>% filter(referenceLimits == 'NO') %>% nrow()` times for reference limits. The full distribution of quantitative length limits and time-to-submit limits is displayed in Figure 2 and limits imposed by individual journals are available in Supplementary Table G1. Table 1 also shows whether PPPR was subject to independent external peer review (for details see Supplementary Table H1).

@@

From Table 2, we can see that more restrictive PPPR options are sometimes accompanied by less restrictive PPPR options. Eleven of the 20 journals only offered one PPPR option. Nine of the 20 journals offered web comments in addition to letters and, in general, web comment policies appeared to be less restrictive than letters. However, this was often unclear because exact quantitative limits were not specified and most differences were marginal. For example, in the JAMA family of journals, letters must be less than 400 words and submitted within 4 weeks of target article publication. By contrast, web comments are marginally less restrictive in terms of length (600 words), and ambiguous about their time-to-submit limits (“We may reject comments because they...are submitted a long time after article publication”). One journal - Science - offered commentaries (called ‘Technical Comments’) in addition to letters and web comments. In this case, letters and commentaries shared a time-to-submit limit of 3 months, and commentaries had a somewhat less restrictive length limit than letters (1000 vs. 300 words). By contrast, no time-to-submit limit was specified for web comments and an ambiguous length limit was implied (web comments should be ‘brief’).

@@

# PPPPR estimates when clinical medicine is excluded
d_freq_noMed <- d_freq_primary %>% filter(ESI_field != "CLINICAL MEDICINE")
linkedPPPR_n_noMed <- d_freq_noMed %>% filter(linkedPPPR == T) %>% nrow()
prev_est_noMed_primary <- percentCI(linkedPPPR_n_noMed, nrow(d_freq_noMed))
```
PPPR prevalence estimates varied substantially across scientific domains (Figure \@ref(fig:fig-practice-prev); for equivalent tabular data see Supplementary Table \@ref(tab:tab-practice-prev)). The number of domains which had zero instances of PPPR was `r primary_zero` according to our primary prevalence estimate and `r secondary_zero` according to our secondary prevalence estimate. The domain with the most PPPR was Clinical Medicine with `r primary_max$primary`% of research articles being linked to PPPR and `r secondary_max$secondary`% of assessed articles being themselves instances of PPPR.

@@

Our exploration of how PPPR is used in practice suggested that letters are far more common than other types of PPPR, perhaps because they are the most frequently available PPPR option and also because, as formal articles, they may impart greater academic credit to their authors than informal web comments (Table 3). We found that half of the PPPRs we examined were behind a paywall, sometimes even when the target article was publicly accessible. This reduces access to PPPR for both professional scientists and other readers, like patients, journalists, and policy-makers (Winker, 2015). Most PPPR had conflict of interest statements and half of those statements declared a potential conflict. Conflict of interest statements enable readers to evaluate an important risk of bias and seem just as relevant to PPPR as they are to other academic articles (Dunn et al., 2016).

The PPPRs we examined addressed a range of issues spanning the timeline of a research project, including design, implementation, analysis, reporting, and interpretation. The vast majority of PPPRs did not include new analysis of original or new data. This may be because the vast majority of original articles did not share data, as is typical in many scientific domains (Hardwicke et al., 2021; Hardwicke, Wallach, et al., 2020; Serghiou et al., 2021). Most of the PPPR was short (~250 words) and published within five months of the target article, perhaps partly because they were published in some journals that imposed the strictest limits on PPPR (Table 2).

@@

The cultivation, documentation, and dissemination of post-publication critique is an important part of a healthy and self-correcting research literature. Our study reveals considerable variation in how PPPR is handled by journals operating across scientific domains. Clinical Medicine had a much more active PPPR culture than other domains; but it's journals also imposed the strictest limits. Overall, PPPR appears to be tightly controlled and restricted by influential academic journals. At many journals, it was not possible to publish PPPR at all, and journals that did offer PPPR options often imposed strict length and time-to-submit restrictions. The PPPR we did identify appeared to have negligible impact on target article authors. These data provide a stratum of empirical evidence upon which to base debates about how scientific critique should be optimally handled. We encourage stakeholders across the academic ecosystem to explore ways to foster a richer culture of post-publication critique.

>>>>

abstract: |
  Critical scrutiny of published scientific research, often called post-publication peer review (PPPR), can highlight errors, limitations, or alternative interpretations that were not adequately addressed during traditional peer review. Currently, there is limited empirical data documenting how academic journals handle this important aspect of scientific self-correction. We assessed PPPR policies and practice amongst 330 influential journals operating across 22 scientific disciplines. 123 (37%) journals did not offer any options for submitting PPPR. 207 (63%) journals offered PPPR options such as letters, commentaries, or web comments, but often imposed limits on length (median 1000, IQR 700 words) and time-to-submit (median 12, IQR 22 weeks). The most restrictive limits were 175 words and 2 weeks. Of a random sample of 2066 research articles published in 2018 by PPPR-offering journals, we found that 1.9% (95% CI [1.4, 2.6]) were the subject of at least one PPPR. Examination of 58 PPPRs found, for example, that 44 received an author reply, of which 41 asserted that the authors’ conclusions were unchanged. Clinical Medicine stood out as the domain with the most active PPPR culture, with all 15 journals providing PPPR options and publishing the most PPPR, but also imposing the strictest length limits (median 400, IQR 150 words) and time-to-submit limits (median 4, IQR 2 weeks). Overall, our findings suggest that influential academic journals often pose serious barriers to the cultivation, documentation, and dissemination of PPPR. We propose several policy changes that journals could adopt to promote a richer culture of post-publication critique.

@@

Table 1 shows how often journal policies imposed limits on PPPR in terms of length, time-to-submit, or number of references. Limits were mostly expressed quantitatively, but sometimes they were qualitative and more ambiguous, for example, stating that PPPR should be ‘concise’ or address ‘recently published’ articles. Often there was no information at all about a particular limit. Occasionally policies explicitly asserted that there was no limit. This happened `r d_policy %>% filter(wordLimits == 'NO') %>% nrow()` time for length limits, `r d_policy %>% filter(timeLimits == 'NO') %>% nrow()` times for time-to-submit limits, and `r d_policy %>% filter(referenceLimits == 'NO') %>% nrow()` times for reference limits. The full distribution of quantitative length limits and time-to-submit limits is displayed in Figure 2 and limits imposed by individual journals are available in Supplementary Table G1. Table 1 also shows whether PPPR was subject to independent external peer review (for details see Supplementary Table H1).

@@

From Table 2, it is notable that 8 of the journals offering the most restrictive PPPR options operate in the domain of Clinical Medicine. Overall in this domain, journals specified strict limits on length (19 PPPR policies with a quantitative limit: median 400, IQR 150 words; 3 PPPR policies did not mention a limit) and time-to-submit (13 PPPR policies with a quantitative limit: median 4, IQR 2 weeks; 3 PPPR policies stated a qualitative limit, 5 PPPR policies did not mention a limit, 1 PPPR policy stated there was no limit). Table 2 also suggests that restrictive PPPR options are sometimes accompanied by less restrictive PPPR options. Eleven of the 20 journals only offered one PPPR option. Nine of the 20 journals offered web comments in addition to letters and, in general, web comment policies appeared to be less restrictive than letters. However, this was often unclear because exact quantitative limits were not specified and most differences were marginal. For example, in the JAMA family of journals, letters must be less than 400 words and submitted within 4 weeks of target article publication. By contrast, web comments are marginally less restrictive in terms of length (600 words), and ambiguous about their time-to-submit limits (“We may reject comments because they...are submitted a long time after article publication”). One journal - Science - offered commentaries (called ‘Technical Comments’) in addition to letters and web comments. In this case, letters and commentaries shared a time-to-submit limit of 3 months, and commentaries had a somewhat less restrictive length limit than letters (1000 vs. 300 words). By contrast, no time-to-submit limit was specified for web comments and an ambiguous length limit was implied (web comments should be ‘brief’).

```{r}
# examine limits in medicine only
med_journ_limits <- d_policy %>% filter(field == "CLINICAL MEDICINE", !is.na(PPPR_name))

med_journ_limits %>% count(wordLimits)
med_journ_limits %>% count(timeLimits)

med_journ_limits %>% select(wordLimits) %>% filter(wordLimits != "NOT STATED") %>% mutate(wordLimits = as.numeric(wordLimits)) %>% pull(wordLimits) %>% median()
med_journ_limits %>% select(wordLimits) %>% filter(wordLimits != "NOT STATED") %>% mutate(wordLimits = as.numeric(wordLimits)) %>% pull(wordLimits) %>% IQR()
med_journ_limits %>% select(timeLimits) %>% filter(timeLimits %notin% c("NOT STATED", "Qualitative limit", "NO")) %>% mutate(timeLimits = as.numeric(timeLimits)) %>% pull(timeLimits) %>% median()
med_journ_limits %>% select(timeLimits) %>% filter(timeLimits %notin% c("NOT STATED", "Qualitative limit", "NO")) %>% mutate(timeLimits = as.numeric(timeLimits)) %>% pull(timeLimits) %>% IQR()
```

@@

# PPPPR estimates when clinical medicine is excluded
d_freq_noMed <- d_freq_primary %>% filter(ESI_field != "CLINICAL MEDICINE")
linkedPPPR_n_noMed <- d_freq_noMed %>% filter(linkedPPPR == T) %>% nrow()
prev_est_noMed_primary <- percentCI(linkedPPPR_n_noMed, nrow(d_freq_noMed))
```

PPPR prevalence estimates varied substantially across scientific domains (Figure \@ref(fig:fig-practice-prev); for equivalent tabular data see Supplementary Table \@ref(tab:tab-practice-prev)). The number of domains which had zero instances of PPPR was `r primary_zero` according to our primary prevalence estimate and `r secondary_zero` according to our secondary prevalence estimate. The domain with the most PPPR was Clinical Medicine with `r primary_max$primary`% of research articles being linked to PPPR and `r secondary_max$secondary`% of assessed articles being themselves instances of PPPR.

@@

Our exploration of how PPPR is used in practice suggested that letters are far more common than other types of PPPR, perhaps because they are the most frequently available PPPR option and also because, as formal articles, they may impart greater academic credit to their authors than informal web comments (Table 3). We found that half of the PPPRs we examined were behind a paywall, sometimes even when the target article was publicly accessible. This reduces access to PPPR for both professional scientists and other readers, like patients, journalists, and policy-makers (Winker, 2015). Most PPPR had conflict of interest statements and a third of those statements declared a potential conflict. Conflict of interest statements enable readers to evaluate an important risk of bias and seem just as relevant to PPPR as they are to other academic articles (Dunn et al., 2016).

The PPPRs we examined addressed a range of issues spanning the timeline of a research project, including design, implementation, analysis, reporting, and interpretation. The vast majority of PPPRs did not include new analysis of original or new data. This may be because very few original articles stated that data was available, as is typical in many scientific domains (Hardwicke et al., 2021; Hardwicke, Wallach, et al., 2020; Serghiou et al., 2021). Most of the PPPR was short (~250 words) and published within five months of the target article, perhaps partly because they were published in some journals that imposed the strictest limits on PPPR (Table 2).

@@

The cultivation, documentation, and dissemination of post-publication critique is an important part of a healthy and self-correcting research literature. Our study reveals considerable variation in how PPPR is handled by journals operating across scientific domains. Clinical Medicine had a much more active PPPR culture than other domains; but it's journals also imposed the strictest limits. Overall, PPPR appears to be tightly controlled and restricted by influential academic journals. At many journals, it was not possible to publish PPPR at all, and journals that did offer PPPR options often imposed strict length and time-to-submit restrictions. The PPPR we did identify appeared to have negligible impact on target article authors' conclusions. These data provide a stratum of empirical evidence upon which to base debates about how scientific critique should be optimally handled. We encourage stakeholders across the academic ecosystem to explore ways to foster a richer culture of post-publication critique.
